---
title: "09-01-inclass"
author: "Emily Beam"
date: "03-30-2021"
output: html_document
---


## Set up

Make sure to run this code once to load our items
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  # For ggplot, dplyr, and friends
library(gapminder)  # For gapminder data
library(scales)     # For nice axis labels
```

## Create a new data set 

Let's make a filtered data set that includes only two years, 2002 and 2007. Note how we save it to an object, `gapminder_filtered`

```{r filtered data}
# Make sure you ran the chunk above to load your libraries

gapminder_filtered <- gapminder %>% 
  filter(year > 2000)
```


## Make a base plot 

```{r base_plot}
## Start by making a facted plot 

base_plot <- ggplot(data = gapminder_filtered,
                    mapping = aes(x = gdpPercap, y = lifeExp, 
                                  color = continent, size = pop)) +
  geom_point() + 
  facet_wrap(vars(year))

base_plot

```

Now, fix the x-axis so it is logged gdpPercap. You will need a type of `scale()` function. Note that if you paste `labels = dollar_format(accuracy = 1)` inside the parentheses of your scale function, you can change how the dollar values are presented!

```{r axis}

base_plot <- base_plot + CODE HERE

base_plot
```

What does this modification do? 
```{r scale_label}

base_plot <- base_plot + 
  
  scale_size_continuous(labels = comma)
  
base_plot
```


Let's adjust the color scale. Oceania is really hard to see! What can we do? We can exclude the top-most value of the range by adding `end = 0.9` to the `scale_color_viridis_d` option. Because it's a new option, make sure to separate by a comma. 

Fix the color scale now:
```{r color_scale}
base_plot <- base_plot + 
  # Use viridis
  scale_color_viridis_d(option = "plasma") 

base_plot




```

Now, let's add some labels: 

Remember the generic form: 

```{}
labs(x = "XTITLE", y = "YTITLE",
       `color = "COLOR LEGEND TITLE", size = "SIZE LEGENDTITLE",
       `title = "TITLE",
       `subtitle = "SUBTITLE",
       caption = "CAPTION)
```

```{r labs}
base_plot <- base_plot + 
  labs(STUFFF)

base_plot
```


## Optional  - play with `ggThemeAssist`

If you're feeling adventurous, you can mess more with ggThemeAssist: 

```{r ggtheme}
#install.packages("ggThemeAssist")
library(ggThemeAssist)

base_plot

# First, run this chunk to load the library 
# Then, highlight base_plot. 
# Then click on "Addins" and select "ggplot Theme Assistant"
# Play! 
# When you're done, check out the code you've created. 
```



## Export
Finally, let's export it as a png file! 

Remember the generic form : 
```{}
ggsave(filename = "NAME.png", plot = TITLE, width = 6, height = 4.5, units = "in")
```


```{r export}

# Export to a png file
```


**Upload your *png file* to blackboard, and you're done!**
