---
title: "10-02 in-class activity"
author: "Emily Beam"
date: "04-08-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("janitor")
# Run me first! 
library(tidyverse)
library(readr)
```

## Getting Started
We're going to dig deeper into the correlates of mobility by using [Opportunity Atlas](https://opportunityatlas.org/) to create some plots. We're going to make a scatter plot!

1. Visit [Opportunity Atlas](https://opportunityatlas.org/) and search for the place you chose for your pre-class activity. If you no longer like this place, you can pick a new one. 

2. Zoom in or out, as necessary, so that you get a "reasonable" geographic area that contains 30-50 census tracts (not a hard and fast rule, but each tract will be one data point).

3. Click on "Download the data" at the bottom right-hand corner. 

  - You want to download "Just what is on the screen." If this option is not available, make sure you select "Single Subgroup." 
  - You can pick "all" or select a specific subgroup, but I wouldn't recommend any group that leads to a lot of grey tracts. 
  - I recommend you stick with "household income" as your outcome variable, but feel free to select a different outcome variable.

4. Get all your options set up and download the data set. 

5. Pick a second variable you want to correlate with your first variable, and download that in the same way. I recommend "Job growth rate from 2004 to 2013" or "Density of Jobs in 2013." You want the first to be an "outcome" and the second to be a "neighborhood characteristic."

**In any case, I'd recommend you pick the same variables as your groupmates, so when you compare graphs it makes sense!**


## Loading data

Make sure your data files are in the right working directory. You may need to adjust the names.
```{r message=FALSE}
# Import outcome 1 and save as a df

income <- read_csv("shown_tract_kfr_rP_gP_pall.csv")

# Import outcome 2 and save as a df

jobdensity <- read_csv("shown_tract_job_density_2013.csv")

```


Take a look at your data (change names if you renamed your data frames!)
```{r}

income
jobdensity
```

Now, we're going to use `join_by()` to merge our two data sets. Note that the `by` part of the function is optional. If you delete it, it will join based on all variables with the same names across the two data sets - in this case, just `tract` and `Name`.

```{r}
merged_data <- full_join(by = c("tract", "Name"), income, jobdensity)
merged_data
```


# Exploratory analysis 

A simple way to look at the relationship between two variables is to calculate a correlation coefficent. Here, I do it using Base R. There are other ways to do this as well. But, this gives us one number, and not a ton of information. 

(Note that you'll need to change the variable names to match yours!)

```{r}

cor.test(merged_data$Household_Income_rP_gP_pall, merged_data$Density_of_Jobs_in_2013,method="pearson")

```


Let's instead make a scatter plot: 

```{r}
  ggplot(data = merged_data, mapping = aes(y = Household_Income_rP_gP_pall, x = Density_of_Jobs_in_2013, color=Name )) + 
      geom_point(show.legend=FALSE) 
  
```


You may want to play around with the scale if you get a lot of bunching. For density of jobs, for example, it looks a lot better if I take a log.
```{r}
  ggplot(data = merged_data, mapping = aes(y = Household_Income_rP_gP_pall, x = Density_of_Jobs_in_2013, color=Name )) + 
      geom_point(show.legend=FALSE) + 
  scale_x_log10()
```

What if you want to estimate a line of best fit? This is regression land, but we can give it a go. Notice here that I moved `color=Name` out of the main ggplot and into `geom_point` only. The idea is that I want to color my points based on city, but I don't want separate regression lines per city. You can play w/ moving arguments around to see how it does or does not make a difference.

```{r}
  ggplot(data = merged_data, mapping = aes(y = Household_Income_rP_gP_pall, x = Density_of_Jobs_in_2013,  )) + 
      geom_point(aes(color=Name), show.legend=FALSE) + 
    geom_smooth(method=lm, se =FALSE, show.legend=FALSE) + 
    scale_x_log10() 
```

# All done!

1. Share w/ your group mates. You can paste in your chat or in your shared Google doc.

2. Knit and upload. If you've got some time, format that graph.
