---
title: 'Lesson 07-02: In-class assignment'
author: "NAME"
output:
  html_document: default
  pdf_document: default
---



## Set-up

This block will install the gapminder package and load it. If this is your first time, you will need to "uncomment" the `install.packages` line in order to install it. You will want to comment it back out or delete it once it's installed, or it can create problems when knitting

```{r load-libraries, warning=FALSE, message=FALSE}
# install.packages("gapminder")
library(tidyverse)
library(gapminder)
```

Now that we've loaded the library, we have access to a *data frame*. Take a look at the data set by running this code. You can also type `View(gapminder)` into the console to see *all* the data, which will show up in a new window.

Note that you can use `head()` and `tail()` to show the first and last six rows of the data as well. What happens when you run `head(gapminder,20)`?

```{r}
# Show a tibble
gapminder

# Show the first six rows (6 is default)
head(gapminder)

# Show the last six rows (6 is default)

tail(gapminder)

## What happens here? 

head(gapminder,20)


```


## Selecting with `select()` 

We can select specific columns or a range of columns using `select()`

What does this do?
```{r select-demo, eval=FALSE, include=FALSE}

#select(gapminder,country)


#select(gapminder,country:lifeExp)

#select(gapminder,-continent)


```

Write two different ways to select country, continent, year, and gdpPercap

```{r select}

# First way to select country, continent, year, and gdpPercal

# Second way to select country, continent, year, and gdpPercal



```
## Filtering with `filter()`

Recall that we can use `filter()` restrict our data according to criteria. We can display it as is, or we can assign it to an object if we want to use that restricted data later. 

You can also use multiple conditions, and these will extract rows that meet every test. By default, if you separate the tests with a comma, R will consider this an “and” test and find rows that are both Denmark and greater than 2000.

What does this do?

```{r filter-demo}

filter(gapminder,gdpPercap < 1000)

# Note that putting parentheses around the entire line below will assign the object /and/ display it

(verypoor <- filter(gapminder, year == 1992, gdpPercap<500))


```

Use `filter()` to do the following: 
```{r filter} 

# Show all data for Canada

## Show all data for Oceania

## Show all rows where the life expectancy is greater than 82

## Show all countries where the life expetancy is greater than 82 and the country is in Africa (you can use $ to select)


```

## Arrange 

We sort our data using `arrange()`

The first line sorts by year, then by country. What does the second line do? What does the third line do? 

```{r arrange-demo}

arrange(gapminder,year, country)


arrange(gapminder,desc(year), continent, country)
```


Use `arrange()` to do the followin:
```{r arrange} 

# Sort the data by year, then by descending life expectancy


```

## Piping
You can nest commands inside other commands to combine them, but your life will be 100% better with... piping! (The keyboard shortcut is `Shift-Command-M` on a Mac, or `Ctrl-Shift-M` on a PC)

```{r piping} 

# Sort by descending life expectancy in 2007

gapminder %>% filter(year == 2007) %>% arrange(desc(lifeExp))

## Show just population data, year and country names when looking at the most populous countries in 1977. Note that you cannot filter on something you didn't select!


## Does not work
#gapminder %>% select(country,pop) %>% filter(year == 1977) %>% arrange(desc(pop))

gapminder %>% select(country,year,pop) %>% filter(year == 1977) %>% arrange(desc(pop))

```

Use `arrange()` and `filter()` to do the following:
```{r arrange-filter} 

# Examine the countries with the lowest life expectancy in 2002


```

In general, you can also use help files, Google, and other resources to find new commands. I wanted to list countries that ever had life expectancies below 35. 
```{r extra} 


gapminder %>%  filter(lifeExp < 35) 



```


This is okay, but there are lots of duplicates! I also don't want to pick just one year, because life expectancies aren't always increasing. What's one way to solve this if you don't know where to start? 

1. Googled what I wanted to do "list distinct observations" (it may take a few tries)
2. Looked for a command that worked with frames. 
3. Ended up in the `dplyr` documentation (surprise!) with the [distinct](https://dplyr.tidyverse.org/reference/distinct.html) command. 
4. Reviewed briefly, but then jumped down to examples to make sure I knew how to apply it.

```{r extra-distinct} 


gapminder %>%  filter(lifeExp < 35) %>% distinct(country)


```


## Basic plotting

Next week we'll do more plotting. For now, use what we know about filters to create a second plot that shows the distribution of per-capita GDP for 2007.

```{r histogram}

## What does this plot do? 
gapminder %>%  filter(year == 1992) %>% ggplot(mapping=aes(y=lifeExp)) + 
  geom_histogram()

## Put revised code that plots a histogram for 2007 that plots per-capita GDP


```

## All done! 
Knit me and upload me to BB. If you knit to an html file, you will need to compress it first. 

