---
title: "Pre-Class Activity 07-02 Template"
author: "Your Name Here"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Pre-class activity (07-02)

Welcome to the second pre-class activity! The instructions are embedded here, as well as on Blackboard.


## Step 1: 
Complete the following 3 primers (*you will need to log-in to your RStudio cloud account again*)

1. [Data Visualization Basics](https://rstudio.cloud/learn/primers/1.1)
2. [Working with tibbles](https://rstudio.cloud/learn/primers/2.1)
3. [Isolating data with `dplyr`](https://rstudio.cloud/learn/primers/2.2)

You'll notice that you can work through examples in the tutorial, run your code, and compare against the solutions. I strongly recommend that you copy and paste your code into a R script or R markdown document as you go, which will give you a record of what you've done. You'll also need to bring in some answers in order to generate your PC document (below)


## Step 2: Editing this markdown document

### Calling `dplyr` and `ggplot2` library

If you haven't already, install the tidyverse by running the codeblock below. You will also need to install `babynames`. You will want to "uncomment" the ones you need by deleting the `#` and running. 

```{r include=FALSE}
  
#install.packages("tidyverse",dependencies=TRUE)
#install.packages("babynames")
```

Note that in the code block, the command `include=FALSE` means that this block will not be included when you knit your file!

Now, run this block of code to load your libraries. See how the three dashes to open and close the code, along with the bracket, indicate a code block. Here, I've named this code block "setup". Note that when you start using `tidyverse` it will automatically load our friends `ggplot2` and `dplyr`!

```{r setup, include=FALSE }
  
library(tidyverse)
library(babynames)

```
If it's not automatically displayed, you can click on the "Console" window to make sure that ran. 

### Results from Data Visualization Basics

In this code block, enter the code you wrote for Exercise 2 of "A code template" in the [Data Visualization Basics](https://rstudio.cloud/learn/primers/1.1) tutorial. Because `mpg` is a pre-loaded dataset, you can work with it on your computer without any additional setup!

Exercise 2: Make a scatterplot of `cty` vs `hwy`.

```{r exercise1}

ggplot(data = mpg) + 
geom_point(aes(x = cty, y = hwy))


```


### Results from Isolating Data 

Use the next code chunk to put your results from "Your Name" in the `%>%` section of [Isolating Data with dplyr](https://rstudio.cloud/learn/primers/2.2)

> In the code chunk below, use a combination of dplyr and ggplot2 functions with %>% to:
    Trim babynames to just the rows that contain your name and your sex
    Trim the result to just the columns that will appear in your graph (not strictly necessary, but useful practice)
    Plot the results as a line graph with year on the x axis and prop on the y axis
    
```{r exercise2}
babynames
babynames %>% 
  filter(name == "Emily", sex == "F") %>% 
  select(year, prop) %>% 
  ggplot() +
    geom_line(aes(x = year, y = prop)) +
    labs(title = "Popularity of the name Emily")

```



<!-- ### Results from Derive Information  -->

<!-- In the final code block, put your code for the two challenges in the "Challenges" of the [Deriving information with `dplyr'`](https://rstudio.cloud/learn/primers/2.3) tutorial. Specifically, the code you need to determines how many distinct boys and girls names achieved a rank of 1 in any year.  -->

<!-- >How many distinct boys names acheived a rank of Number 1 in any year? -->
<!-- > How many distinct girls names acheived a rank of Number 1 in any year? -->

<!-- ```{r} -->
<!-- # Insert your code below. Then run it!  -->
<!-- babynames %>%  -->
<!--   group_by(year, sex) %>%  -->
<!--   mutate(rank = min_rank(desc(n))) %>%  -->
<!--   filter(rank == 1, sex == "M") %>%  -->
<!--   ungroup() %>%  -->
<!--   summarise(distinct = n_distinct(name)) -->

<!-- babynames %>%  -->
<!--   group_by(year, sex) %>%  -->
<!--   mutate(rank = min_rank(desc(n))) %>%  -->
<!--   filter(rank == 1, sex == "F") %>%  -->
<!--   ungroup() %>%  -->
<!--   summarise(distinct = n_distinct(name)) -->
<!-- ``` -->

  
### Reflection

In this space, answer the following prompts (no word limit): 

- What are 2 questions you had or concepts you felt uncomfortable with?

- Were there any resources you used to get things working besides the tutorials themselves? 

### Knit! 

Finally, knit this document to Word, html, or PDF and upload into Blackboard. 

If you want to do PDF but are getting an error, install the `tinytex` package.

*Note:* If you get a weird error message about `Error in contrib.url(repos,"source")...` comment out the two lines that install your packages.

